Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
add side stuff

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$
      \end{cases*}$\\
  \end{tabular}
\end{tabular}
\end{centermath}
The \OrLens lens deals with data that can come in one form or another. If the
data gets changed from one format to the other, information in the old format is
lost.
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\ldots\String_n$
%    & =
%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
%    
%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
%    & =
%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
%    
%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
%    & =
%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
%        \Lens.\CreateROf{\String_i} & otherwise
%      \end{cases*}$\\
%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
%    & =
%    & $\String_1'\ldots\String_n'$ where $\String_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
%        \Lens.\CreateROf{\StringAlt_i} & otherwise
%      \end{cases*}$
%  \end{tabular}
%\end{center}
\begin{centermath}
\begin{tabular}[b]{l@{\qquad}l}
$
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt
  }
  {
    \MergeROf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\BRegex_1}{\BRegex_2}
    \Leftrightarrow
    \BRegexAlt
  }
$
&
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\

@@

\begin{centermath}
\begin{tabular}[b]{l@{\qquad}l}
$
\inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
  }
  {
    \IterateLensOf{\Lens} \OfType
    \StarOf{\BRegex}
    \Leftrightarrow
    \StarOf{\BRegexAlt}
  }
  $
  &
  $
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
  }
  {
    \InvertOf{\Lens} \OfType \BRegexAlt \Leftrightarrow \BRegex
  }
  $
\end{tabular}
\end{centermath}
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
The \IterateLens lens deals with iterated data, while inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa. The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.
\[
  \centering
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
    \BRegex \SSREquiv \BRegex'\\
    \BRegexAlt \SSREquiv \BRegexAlt'
  }
  {
    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
  }
\]

Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a
lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is
equivalent to $T'$. Type equivalence is useful both for addressing type
annotations, and for making well-typed compositions.

>>>>

    $\PutL{} \App \StringAlt \App \String$
    & =
    & $\begin{cases*}
        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$\\
        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$
      \end{cases*}$\\
  \end{tabular}
\end{tabular}
\end{centermath}
The \OrLens lens deals with data that can come in one form or another. If the
data gets changed from one format to the other, information in the old format is
lost.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
  }
  {
    \IterateLensOf{\Lens} \OfType
    \StarOf{\BRegex}
    \Leftrightarrow
    \StarOf{\BRegexAlt}
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \String_1\ldots\String_n$
%    & =
%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
%    
%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
%    & =
%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
%    
%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
%    & =
%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
%        \Lens.\CreateROf{\String_i} & otherwise
%      \end{cases*}$\\
%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
%    & =
%    & $\String_1'\ldots\String_n'$ where $\String_i' =
%      \begin{cases*}
%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
%        \Lens.\CreateROf{\StringAlt_i} & otherwise
%      \end{cases*}$
%  \end{tabular}
%\end{center}
The \IterateLens lens deals with iterated data.

\begin{centermath}
\begin{tabular}[b]{l@{\qquad}l}
$
  \centering
  \inferrule*
  {
    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt\\
    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt
  }
  {
    \MergeROf{\Lens_1}{\Lens_2} \OfType
    \RegexOr{\BRegex_1}{\BRegex_2}
    \Leftrightarrow
    \BRegexAlt
  }
$
&
  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
    $\CreateR{} \App \String$
    & =
    & $\begin{cases*}
      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
      \end{cases*}$\\

@@

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
  }
  {
    \InvertOf{\Lens} \OfType \BRegexAlt \Leftrightarrow \BRegex
  }
\]
%\begin{center}
%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
%  \end{tabular}
%\end{center}
Inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa. The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.

\[
  \centering
  \inferrule*
  {
    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
    \BRegex \SSREquiv \BRegex'\\
    \BRegexAlt \SSREquiv \BRegexAlt'
  }
  {
    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
  }
\]
Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a
lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is
equivalent to $T'$. Type equivalence is useful both for addressing type
annotations, and for making well-typed compositions.
